---
title: "Day 2 - Password Philosophy"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Background

Source of the daily challenge comes from the [Advent of Code for 2020](https://adventofcode.com/) produced by [Eric Wastl](https://twitter.com/ericwastl).

## Data Source

The source of data in the expense report comes from: https://adventofcode.com/2020/day/2/input

```{r}
library(dplyr)
library(stringi)
```

## Data management

Read in data from the [advent of code website](https://adventofcode.com/2020/day/2/input):

```{r}
d <- read.csv('Data/input_Day2.txt', header = F, col.names = 'password', colClasses = 'character')
head(d)
```

## Part 1

Determine the number of valid passwords in the North Pole Toboggan Rental Shop database. Each row of the input dataset lists the required passkey for a password to be valid followed by a password. The passkey is defined as a numeric range of instances a a specified character must appear in the password.

## Analysis

### Find the number of all valid passwords in the database

Based on a quick view of the first six rows in the dataframe, it looks like each row has a uniform pattern. The numeric range is separated by a "-", the character of interest is separated from the numeric range by a space, and the password occurs after a ": ". These patterns can be used to extract each of these components.

```{r}
cd <- d %>%
  tidyr::separate(., password, into = c("numeric_range", "character", "password"), sep = " ") %>%
  tidyr::separate(., numeric_range, into = c("numeric_range_min", "numeric_range_max"), sep = "-")

#Transform 'numeric_range_min' and 'numeric_range_max' to class numeric
cd$numeric_range_min <- as.numeric(cd$numeric_range_min)
cd$numeric_range_max <- as.numeric(cd$numeric_range_max)

#Remove ":" from character column
cd$character <- gsub(":", "", cd$character)

#Find the number of times the character of interest is in the password using the stringi packages 'stri_count_regex' function
cd$character_n <- as.numeric(stri_count_regex(cd$password, cd$character))

#Find if the number of times the character of interest is in the password is between the range of accepted numeric values for a valid password
cd$password_valid <- cd$character_n >= cd$numeric_range_min & cd$character_n <= cd$numeric_range_max

#Summarize the count of valid and non-valid passwords
cd %>%
  group_by(password_valid) %>%
  summarise(n = n())

```
The number of valid passwords is 572.

## Part 2

Passwords in the North Pole Toboggan Rental Shop database are valid if the character of interest is in one and only one of the two numeric values listed. The objective is to find the number of passwords that are valid based on these scheme.

## Analysis


```{r}
cd$password_valid2 <- xor(substr(cd$password, cd$numeric_range_min, cd$numeric_range_min) == cd$character, substr(cd$password, cd$numeric_range_max, cd$numeric_range_max) == cd$character)

#Summarize the count of valid and non-valid passwords
cd %>%
  group_by(password_valid2) %>%
  summarise(n = n())
```

The number of valid passwords is 306.