---
title: "Day 1 - Report Repair"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Background

Source of the daily challenge comes from the [Advent of Code for 2020](https://adventofcode.com/) produced by [Eric Wastl](https://twitter.com/ericwastl).

## Data Source

The source of data in the expense report comes from: https://adventofcode.com/2020/day/1/input

```{r}
library(dplyr)
```

## Data management

Read in data from the [advent of code website](https://adventofcode.com/2020/day/1/input):

```{r}
d <- read.csv('Data/input_Day1.txt', header = F, col.names = 'expense', colClasses = 'numeric')
head(d)
```

## Part 1

Find two numbers in the expense report that sum to 2020 and multiply the two numbers together.

## Analysis

### Find two numbers that add to 2020
Print summary statistics for the dataset

```{r}
summary(d)
```

Based on the minimum value of 48, we can remove all expenses greater than 1972 since the sum would exceed 2020.

```{r}
cd <- d %>%
  filter(expense <= 1972)

```

Loop through possible expense summations to see which two expenses add to 2020.

```{r}
exp_list <- sort(cd$expense)

i <- j <- c()

for(i in exp_list){
  j <- 2020-i
  if(j %in% exp_list){
    print(c(i,j))
    break()
  }
}

```

Expenses of 251 and 1769 add to 2020. 

### Multiply these two numbers together

```{r}
exp_mult <- i*j
```

The final result is:
```{r}
print(exp_mult)
```

## Part 2

Find three numbers in the expense report that sum to 2020 and multiply the three numbers together.

## Analysis

### Find three numbers that add to 2020
Loop through possible expense summations to see which three expenses add to 2020.

```{r}
i <- j <- k <- c()

for(i in exp_list){
  exp_list2 <- exp_list[exp_list != i]
  for(j in exp_list2){
    k <- 2020 - (i + j)
    if(k %in% exp_list2){
      print(c(i,j,k)) 
      break()
    }
  }
break()}
```

Expenses of 48, 383, and 1589 add to 2020. 

### Multiply these three numbers together

```{r}
exp_mult <- i*j*k
```

The final result is:
```{r}
print(exp_mult)
```
